Realizability-Preserving DG-IMEX Method for the Two-Moment Model of Fermion Transport \tnoteref{support}

### 2018-04-12 Documents/M1/realizableFermionicM1/sections/imex.tex
Added comment on PARSD

We therefore proceed to design GSA positivity-preserving IMEX schemes, without the correction step, that perform better in the diffusion limit.  
To this end we define the vectors $\vec{\cJ}=(\cJ^{(1)},\ldots,\cJ^{(s)})^{T}$ and $\vec{\vect{\cH}}=(\vect{\cH}^{(1)},\ldots,\vect{\cH}^{(s)})^{T}$.  
In the context of IMEX schemes, the diffusion limit balance implies that the relation $A\,\vec{\vect{\cH}}=-\f{1}{3}\,\tau\,\tilde{A}\,\nabla\vec{\cJ}$ should hold to relate the stage values in Eq.~\eqref{imexStages}.  
For the final stage ($i=s$), 
\begin{equation}
  \vect{\cH}^{(s)}=-\f{1}{3}\,\tau\,\vect{e}_{s}^{T}A^{-1}\tilde{A}\vect{e}\,\nabla\cJ^{n}+\cO(\dt\,\tau),
\end{equation}
where $\vect{e}_{s}=(0,\ldots,0,1)^{T}$ and $\vect{e}=(1,\ldots,1)^{T}$.  
Thus, for $\vect{\cH}^{(s)}$ to be accurate in the diffusion limit, we require the IMEX coefficients to satisfy
\begin{equation}
  \vect{e}_{s}^{T}A^{-1}\tilde{A}\vect{e} = 1.
  \label{eq:diffusionCondition}
\end{equation}
Unfortunately, this requirement, together with the order conditions given by Eqs.~\eqref{orderConditions1} and \eqref{orderConditions2} (with $\alpha=0$), and the positivity conditions on $c_{ij}$ and $\tilde{c}_{ij}$ results in too many constraints.  
(\ee{Can we prove that it is impossible to find a second-order accurate, positivity-preserving IMEX scheme that is accurate in the diffusion limit?  Does the answer to this depend on $s$?})
(We are concerned about increasing the number of stages, and implicit solves, since the implicit solve will dominate the computational cost of the IMEX scheme with a more realistic collision operator.)
To reduce the number of constraints, and accommodate accuracy in the diffusion limit, we relax the requirement of overall second-order accuracy of the IMEX scheme.  
Instead, we only require the scheme to be second-order accurate in the streaming limit ($\vect{\cQ}=0$).  
This gives the order conditions
\begin{equation}
  \sum_{i=1}^{s}\tilde{w}_{i}=1
  \quad\text{and}\quad
  \sum_{i=1}^{s}\tilde{w}_{i}\,\tilde{c}_{i}=\f{1}{2}.  
  \label{eq:orderConditionsEx}
\end{equation}
We then seek to design a GSA IMEX scheme with the following properties
\begin{itemize}
  \item Second-order in the streaming limit; i.e., satisfies Eq.~\eqref{eq:orderConditionsEx}.
  \item Well-behaved in the diffusion limit; i.e., satisfies Eq.~\eqref{eq:diffusionCondition}.
  \item Positivity-preserving (conditions given by \cite{hu_etal_2017}; with $c_{\mbox{\tiny Sch}}$ maximal).
  \item Few implicit stages.
\end{itemize}
Fortunately, IMEX schemes satisfying these properties are easy to find, and we provide an example of type ARS with $s=3$ (two implicit solves) in \ref{app:butcherTables} (scheme PARSD).  
This scheme is similar to the scheme given in \cite{mcclareen_2008} (see scheme PC2 in \ref{app:butcherTables}), which is also a GSA IMEX scheme of type ARS with $s=3$.  
Scheme PC2 is second-order in the streaming limit, has been demonstrated to work well in the diffusion limit \cite{mcclareen_2008,radice_etal_2013}, and satisfies the positivity conditions, but $c_{\mbox{\tiny Sch}}=0$ (our primary motivation for finding an alternative).  
In Section~\ref{sec:numerical}, we show numerically that the accuracy of scheme PARSD is practically identical to the accuracy of scheme PC2.

>>>>

We therefore proceed to design GSA positivity-preserving IMEX schemes, without the correction step, that perform better in the diffusion limit.  
To this end we define the vectors $\vec{\cJ}=(\cJ^{(1)},\ldots,\cJ^{(s)})^{T}$ and $\vec{\vect{\cH}}=(\vect{\cH}^{(1)},\ldots,\vect{\cH}^{(s)})^{T}$.  
In the context of IMEX schemes, the diffusion limit balance implies that the relation $A\,\vec{\vect{\cH}}=-\f{1}{3}\,\tau\,\tilde{A}\,\nabla\vec{\cJ}$ should hold to relate the stage values in Eq.~\eqref{imexStages}.  
For the final stage ($i=s$), 
\begin{equation}
  \vect{\cH}^{(s)}=-\f{1}{3}\,\tau\,\vect{e}_{s}^{T}A^{-1}\tilde{A}\vect{e}\,\nabla\cJ^{n}+\cO(\dt\,\tau),
\end{equation}
where $\vect{e}_{s}=(0,\ldots,0,1)^{T}$ and $\vect{e}=(1,\ldots,1)^{T}$.  
Thus, for $\vect{\cH}^{(s)}$ to be accurate in the diffusion limit, we require the IMEX coefficients to satisfy
\begin{equation}
  \vect{e}_{s}^{T}A^{-1}\tilde{A}\vect{e} = 1.
  \label{eq:diffusionCondition}
\end{equation}
Unfortunately, this requirement, together with the order conditions given by Eqs.~\eqref{orderConditions1} and \eqref{orderConditions2} (with $\alpha=0$), and the positivity conditions on $c_{ij}$ and $\tilde{c}_{ij}$ results in too many constraints.  
(\ee{Can we prove that it is impossible to find a second-order accurate, positivity-preserving IMEX scheme that is accurate in the diffusion limit?  Does the answer to this depend on $s$?})
(We are concerned about increasing the number of stages, and implicit solves, since the implicit solve will dominate the computational cost of the IMEX scheme with a more realistic collision operator.)
To reduce the number of constraints, and accommodate accuracy in the diffusion limit, we relax the requirement of overall second-order accuracy of the IMEX scheme.  
Instead, we only require the scheme to be second-order accurate in the streaming limit ($\vect{\cQ}=0$).  
This gives the order conditions
\begin{equation}
  \sum_{i=1}^{s}\tilde{w}_{i}=1
  \quad\text{and}\quad
  \sum_{i=1}^{s}\tilde{w}_{i}\,\tilde{c}_{i}=\f{1}{2}.  
  \label{eq:orderConditionsEx}
\end{equation}
We then seek to design a GSA IMEX scheme with the following properties
\begin{itemize}
  \item Second-order in the streaming limit; i.e., satisfies Eq.~\eqref{eq:orderConditionsEx}.
  \item Well-behaved in the diffusion limit; i.e., satisfies Eq.~\eqref{eq:diffusionCondition}.
  \item Positivity-preserving (conditions given by \cite{hu_etal_2017}; with $c_{\mbox{\tiny Sch}}$ maximal).
  \item Few implicit stages.
\end{itemize}
Fortunately, IMEX schemes satisfying these properties are easy to find, and we provide an example of type ARS with $s=3$ (two implicit solves) in \ref{app:butcherTables} (scheme PARSD).  
In the streaming limit, this scheme is identical to the optimal second-order accurate strong-stability preserving Runge-Kutta method \cite{gottlieb_etal_2001}.  
It is also very similar to the scheme given in \cite{mcclareen_2008} (see scheme PC2 in \ref{app:butcherTables}), which is also a GSA IMEX scheme of type ARS with $s=3$.  
Scheme PC2 is second-order in the streaming limit, has been demonstrated to work well in the diffusion limit \cite{mcclareen_2008,radice_etal_2013}, and satisfies the positivity conditions, but $c_{\mbox{\tiny Sch}}=0$ (our primary motivation for finding an alternative).  
In Section~\ref{sec:numerical}, we show numerically that the accuracy of scheme PARSD is practically identical to the accuracy of scheme PC2.
